---
title: "Introduction to R."
author: "Macià Salvà Salvà"
date: "21746 Data Mining 2022-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data processing

First, Inlude the dataset:

```{r}
df=read.table(file="auto-mpg.data")
head(df)
```

The data set vars has no name, we should put name to the columns: 

```{r}
colnames(df)=c("mpg","cylinders","displacement","horsepower","weigth","acceleration","model year","origin","car name")
head(df)
```

Look the characteristics of dataset:

```{r}
str(df)
```

In this data frame, the horsepower is "chr" values and it must be double or integer. Also in the dataset, are null values and they must be removed.


```{r}
df$horsepower <-as.numeric(as.character(df$horsepower))
df=na.omit(df)
str(df)
```

Let's print a summary of our data:

```{r}
name=colnames(df)
par(mfrow = c(2, 4))
for(i in 1:length(name)) {
  if(i!=length(name)){
    plot(df[, name[i]], main=toupper(name[i]), ylab=name[i],col="#33D4FF") 
  }
}
```

We can see that cylinders 3 and 5 are not good data. We have very few examples. Let's see what examples we have.

```{r}
#This function print the sum of cylinders examples
printSumCylinders <- function() {
  v <- c(sum(df$cylinders=="3"),sum(df$cylinders=="4"),sum(df$cylinders=="5"),sum(df$cylinders=="8"))

  n <- c("3","4", "5", "8")
  m <- data.frame("Cylynders"=n,"Valores"=v)
  print(m)
}
```


```{r}
printSumCylinders()
```

We can depreciate the valors of cylinders 3 and 5, because are not good examples. Also between 4 and 8 cylinders is not the best data because we have the double on 4 Cylinders with respect to the 8 cylinders.

```{r}
df=df[df$cylinders!=3, ]
df=df[df$cylinders!=5, ]

printSumCylinders()
```
```{r}
head(df)
```

## 2. Questions

### 2.1 What is the relation between Horsepower and the cylinders?

A priori, we think that more cylinders in a car are equivalent to say, this car has more power.
Let's make a graph between these two variables.

```{r}
boxplot(df$horsepower~df$cylinders, pch=4,main="Horsepower vs Cylinders", xlab="Horsepower", ylab="Cylinders", col='#FF5733')
```

We can see that it is true because is logical when we have more horsepower have more cylinders.

```{r}
aggregate(horsepower~cylinders,df,mean)
```

```{r}
cy=df$cylinders
hp=df$horsepower
cov(df$cylinders,df$horsepower)
```

### 2.2 The trend of each year is to have cars with more power?

This could be an interesting question to ask. But, with these data set can really answer that? Do people normally have sports cars? Definitely not.  

But let's see what can be found in the data set. For this question, we decide that if a car has more than 120 horsepower is a power car.

```{r}
plot(df$`model year`,df$horsepower, main="Cylinders with Weigth", xlab="Horsepower", ylab="Model year",pch=20, col= ifelse(df$horsepower > 120, 'red','green'))
```

With these data this question cannot be solved. We also do not take into account hybrid cars and electric cars. Furthermore we are looking at older cars.

### 2.3 Which cylenders has the better MPG ratio?

I think that if we have more cylinders, we will have less relation MPG. Because we have seen that, if we have more cylinders, we have more power. That power will need more gallons and will produce less relation between MPG.

In these page, we can learn more about MPG: https://reviews.tn/es/wiki/what-is-a-bad-mpg/

Now we will consider:

* If MPG > 35 good relation.
* If 25 < MPG < 35 is not the best relation.
* If MPG < 25 is a bad relation.

```{r}
plot(df$cylinders,df$mpg,pch=18,main="MPG vs Cylenders", xlab="MPG", ylab="Cylinders", col = ifelse(df$mpg >35,'green',ifelse(df$mpg >25,'yellow','red')))
```

Now, we can see that the 4 cylinders have the best MPG ratio. It is obvious because they are the least powerful cars in our data set.


### 2.4 What is the relation between the displacement and the cylenders?

Of course. If we search the formula of de displacemnet (https://es.wikipedia.org/wiki/Cilindrada#:~:text=Cilindrada%20es%20la%20denominaci%C3%B3n%20que,equivalen%20a%20una%20pulgada%20c%C3%BAbica.), if we have more cylinders, we will have more displacement.
 
Formula:
$$UnitDisplacement = \frac{\Pi * D^{2}}4 * L $$
$$TotalDisplacement = \frac{\Pi * D^{2}}4 * L *n$$

Where :

* D is the diameter of the cylinder.
 
* L is the longitud of the piston.

* n totals of the cylinders.

```{r}
plot(df$cylinders,df$displacement,main="Displacement vs Cylenders", xlab="Cylinders", ylab="Displacement",col = ifelse(df$displacement >300,'red',ifelse(df$displacement >200,'yellow','green')))
```

The red colors are the cars with the largest displacement, while the green colors are the cars with the smallest displacement.

We can see the covariance that good.

```{r}
cov(df$displacement,df$cylinders)
```

It is logic see this relation, also we have deduce it on the intro of these question, and we can see this dependence on the formula

### 2.5 The more cylinders a car has, the heavier it tends to weigth?

All the vehicles that are in this data set are street vehicles. The vehicles of competition have more cylinders and their weight is its weight is very small.

But in the street, the more powerful vehicles have bigger engines. This affects the overall weight of the vehicle.


```{r}
plot(df$cylinders,df$weigth,main="Cylinders with Weigth", xlab="Weigth", ylab="Cylinders", pch=18,col= ifelse(df$cylinders==4,'#DEB887',ifelse(df$cylinders==6,'#5F9EA0','#FF7F50')))
```

We can see a relation between weight and the cylinders. When more cylinders have the car, more weight is it.

## 3. Data Distribution

With a corrplot we can see the dependencies of our variables. 

```{r}
library(corrplot)
corrplot(cor(df[1:8]),method= "number") # colorful number
```

Some of the questions answered, we can see their relationships by means of this matrix. For example between the MPG and the cylinders variable we found a -0.79 is close to 1, that means that there is a relation between these variables.

Also, we can see some variables that not have connexions between them. For example, the  'Model year' variable between 'acceleration'  variable the relationship is not close. But it is logical since they have nothing to do with each other. In a year, cars can been created with higher and lower acceleration.


Let's see the distribution:

```{r}
name=colnames(df)
par(mfrow = c(2, 4))
for(i in 1:length(name)) {
  if(i!=length(name)){
    hist(df[, name[i]], main=toupper(name[i]),ylab=name[i],xlab = "distribution",col="#33D4FF") 
  }
}
```

The distribution of our data is correct and logical. For example we find values such as horsepower, ranging from 50 to 150. This distribution is logical since a normal car has at least 50 horsepower.
