---
title: "Linear Regression 1"
author: "Margaret Miró"
output: 
  html_document: 
    toc: yes
  pdf_document:
    keep_tex: true
---

he de fer lm(Y - X, )

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## The Data: trees

In this example we will be using the trees dataset available in R.
```{r trees}
summary(trees)
```

You can also embed plots:
```{r volume and height, echo=FALSE}
plot(trees$Volume~trees$Height,main="Timber volume vs. Tree height",
     xlab="height (ft)",ylab="volume (sq ft)")
```

## Linear Regression with lm().

The regression parameters can be calculated using the "lm()" (linear model) function that estimates the slope and y-intercept of the fitted line.
```{r}
trees.lm<-lm(trees$Volume~trees$Height)
trees.lm
```

### Function for Predictions and Errors. 
Let's check if the regression line passes through the point of means $(\overline{x},\overline{y})$, that is, the predicted value for $\overline{x}$, $\beta_0 + \beta_1 \cdot \overline{x}$ and $\overline{y}$ are equal. We start by creating a function "Yhat" (predicted value).

```{r}
Yhat=function(x,beta0,beta1){
 beta0+beta1*x
}
Yhat(mean(trees$Height),trees.lm$coefficient[1],trees.lm$coefficient[2])
mean(trees$Volume)
```

We have obtained the same results. This, the regression line passes through $(\overline{x},\overline{y})$.

What can you say about the mean of the errors? Calculate it.
```{r}
mean(trees.lm$residuals)
```

### Determination coefficient.

We will find the determination coefficient, adjusted r-squared using the "lm" output. No es muy b    ueno.
```{r}
### Using lm" output
summary(trees.lm)$r.squared
```

## Volume as function of Girth

The results for the linear model Volume as a function of Height are not very good. Let's study Volume as a function of Girth.
```{r}
trees.lm2<-lm(trees$Volume~trees$Girth)
summary(trees.lm2)
```



